---
title: "Covid-19 Situation in Malaysia"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    social: menu
    source_code: https://github.com/wnarifin/covid-19-malaysia/
    navbar:
      - {title: "SIR Model", href: "https://wnarifin.github.io/covid-19-malaysia-sir/", align: right}
    theme: default
---

```{r setup, include=FALSE}
library(flexdashboard)
library(ggplot2)
library(plotly)
library(broom)

# --- Malaysia ---
# covid_my_full = read.csv("https://wnarifin.github.io/covid-19-malaysia/covid-19_my_full.csv")
covid_my_full = read.csv("covid-19_my_full.csv")
covid_my_full$date = as.Date(covid_my_full$date)

# --- state ---
# covid_my_state = read.csv("https://wnarifin.github.io/covid-19-malaysia/covid-19_my_state.csv")
covid_my_state = read.csv("covid-19_my_state.csv")
covid_my_state$date = as.Date(covid_my_state$date)
data_plots = covid_my_state

sel_location = "Malaysia"
sel_date = max(covid_my_full$date)

date_now = max(covid_my_full$date)
days_first_now = max(covid_my_full$days_first)
days_tabligh_now = max(covid_my_full$days_tabligh)
days_mco_now = max(covid_my_full$days_mco)
days_cmco_now = max(covid_my_full$days_cmco)
days_rmco_now = max(covid_my_full$days_rmco)

# cases
model1 = glm(new_cases ~ days_tabligh + days_mco + days_cmco + days_rmco, family = "quasipoisson",  data = covid_my_full)
max_day = 7
data_week = data.frame(date = date_now + 1:max_day, new_cases = rep(NA, max_day), total_cases = rep(NA, max_day),
                       days_tabligh = days_tabligh_now + 1:max_day, days_mco = days_mco_now + 1:max_day,
                       days_cmco = days_cmco_now + 1:max_day, days_rmco = days_rmco_now + 1:max_day)
data_plot = subset(covid_my_full, select = c(date, new_cases, total_cases, days_tabligh, days_mco, days_cmco, days_rmco))
data_plot = rbind(data_plot, data_week)
data_plot$new_cases_predicted = round(predict(model1, data_plot, type = "response"))
data_plot$total_cases_predicted = cumsum(data_plot$new_cases_predicted)

# based on predicted total counts in the past 7 days + future 7 days
data_double = data_plot[tail(which(data_plot$date >= date_now - 7), 14),]
data_double$days = 1:14
model_ = glm(total_cases_predicted ~ days, family = "quasipoisson", data_double)
# log(2) / log(exp(coef(model_)[2]))
days_to_double = log(2) / coef(model_)[2]

model1_rr = tidy(model1, conf.int = T, exponentiate = T)

# deaths
model2 = glm(new_deaths ~ icu, family = "quasipoisson", data = covid_my_full)
# icu to predict new_deaths

# predict icu count, better linear model
model_i = glm(icu ~ new_cases, data = covid_my_full)

# have to use icu_predicted + 7 days
icu_ = round(predict(model_i, list(new_cases = tail(data_plot$new_cases_predicted, 7)), "response"))
data_week1 = data.frame(date = date_now + 1:7, new_deaths = rep(NA, 7), total_deaths = rep(NA, 7), icu = icu_)
data_plot1 = subset(covid_my_full, select = c(date, new_deaths, total_deaths, icu))
data_plot1 = rbind(data_plot1, data_week1)
data_plot1$new_deaths_predicted = round(predict(model2, data_plot1, type = "response"))
data_plot1$total_deaths_predicted = cumsum(data_plot1$new_deaths_predicted)

model2_rr = tidy(model2, conf.int = T, exponentiate = T)
```
by `r paste0(format(sel_date, "%d %B %Y"), ", 12PM")`.

Overview
========

Row
---

### New cases {.value-box}
```{r}
valueBox(
  value = covid_my_full[dim(covid_my_full)[1], "new_cases"],
  icon = "fa-area-chart",
  color = "orange"
)
```

### Active cases {.value-box}
```{r}
valueBox(
  value = covid_my_full[dim(covid_my_full)[1], "total_cases"] - covid_my_full[dim(covid_my_full)[1], "total_recover"] - covid_my_full[dim(covid_my_full)[1], "total_deaths"],
  icon = "fa-area-chart",
  color = "orange"
)
```

### Total cases {.value-box}
```{r}
valueBox(
  value = covid_my_full[dim(covid_my_full)[1], "total_cases"],
  icon = "fa-area-chart",
  color = "orange"
)
```

### cases per 100,000 persons {.value-box}
```{r}
# Malaysian population is 32.68 mil at 4th quarter 2019
valueBox(
  value = round(covid_my_full[dim(covid_my_full)[1], "total_cases"] / (32.68*1000000) * 100000, 2),
  icon = "fa-area-chart",
  color = "orange"
)
```

### days to double the number of total cases {.value-box}
```{r}
valueBox(
  value = round(days_to_double, 2),
  icon = "fa-area-chart",
  color = "orange"
)
```

Row
---

### New deaths {.value-box}
```{r}
valueBox(
  value = covid_my_full[dim(covid_my_full)[1], "new_deaths"],
  icon = "fa-area-chart",
  color = "red"
)
```

### Total deaths {.value-box}
```{r}
valueBox(
  value = covid_my_full[dim(covid_my_full)[1], "total_deaths"],
  icon = "fa-area-chart",
  color = "red"
)
```

### deaths per 1000 cases {.value-box}
```{r}
valueBox(
  value = round(covid_my_full[dim(covid_my_full)[1], "total_deaths"] / covid_my_full[dim(covid_my_full)[1], "total_cases"] * 1000, 2),
  icon = "fa-area-chart",
  color = "red"
)
```

Row
---

### in ICU {.value-box}
```{r}
valueBox(
  value = covid_my_full[dim(covid_my_full)[1], "icu"],
  icon = "fa-area-chart",
  color = "blue"
)
```

### require breathing support {.value-box}
```{r}
valueBox(
  value = covid_my_full[dim(covid_my_full)[1], "support"],
  icon = "fa-area-chart",
  color = "blue"
)
```

Row
---

### Recovered {.value-box}
```{r}
valueBox(
  value = covid_my_full[dim(covid_my_full)[1], "recover"],
  icon = "fa-area-chart",
  color = "green"
)
```

### Total recovered {.value-box}
```{r}
valueBox(
  value = covid_my_full[dim(covid_my_full)[1], "total_recover"],
  icon = "fa-area-chart",
  color = "green"
)
```

### recovered per 1000 cases {.value-box}
```{r}
valueBox(
  value = round(covid_my_full[dim(covid_my_full)[1], "total_recover"] / covid_my_full[dim(covid_my_full)[1], "total_cases"] * 1000, 2),
  icon = "fa-area-chart",
  color = "green"
)
```

Row
---

### **Sources:** Director-General of Health Malaysia and Department of Statistics Malaysia. {.value-box}
```{r}
valueBox(
  value = "",
)
```

Cases
=====

Row
---

### new cases today  {.value-box}
```{r}
valueBox(
  value = data_plot[data_plot$date == date_now, "new_cases"],
  #icon = "fa-area-chart",
  color = "orange"
)
```

### new cases expected tomorrow  {.value-box}
```{r}
valueBox(
  value = data_plot[data_plot$date == date_now + 1, "new_cases_predicted"],
  #icon = "fa-area-chart",
  color = "orange"
)
```

### additional new cases expected in 3 days  {.value-box}
```{r}
valueBox(
  value = data_plot[data_plot$date == date_now + 3, "total_cases_predicted"] - data_plot[data_plot$date == date_now, "total_cases_predicted"],
  #icon = "fa-area-chart",
  color = "orange"
)
```

### additional new cases expected in a week  {.value-box}
```{r}
valueBox(
  value = data_plot[data_plot$date == date_now + 7, "total_cases_predicted"] - data_plot[data_plot$date == date_now, "total_cases_predicted"],
  #icon = "fa-area-chart",
  color = "orange"
)
```

### total cases today  {.value-box}
```{r}
valueBox(
  value = data_plot[data_plot$date == date_now, "total_cases"],
  #icon = "fa-area-chart",
  color = "orange"
)
```

### total cases expected in a week  {.value-box}
```{r}
valueBox(
  value = data_plot[data_plot$date == date_now + 7, "total_cases_predicted"] - data_plot[data_plot$date == date_now, "total_cases_predicted"] + data_plot[data_plot$date == date_now, "total_cases"],
  #value = data_plot[data_plot$date == date_now + 7, "total_cases_predicted"],
  #icon = "fa-area-chart",
  color = "orange"
)
```

Row
---

### New cases, hover/tap on the graph for details.
```{r}
# daily
plot1 = ggplot(data_plot) +
  geom_col(aes(x = date, y = new_cases), fill = "red", alpha = 0.7) +
  geom_line(aes(x = date, y = new_cases_predicted), color = "blue", alpha = 0.7) +
  scale_x_date(date_breaks = "28 day", date_labels = "%d/%m/%y") +
  geom_vline(xintercept = as.numeric(as.Date("2020-02-28")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-02-28")+2, y = max(data_plot$new_cases, na.rm=T)-75, label = "Tabligh", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-03-18")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-03-18")+2, y = max(data_plot$new_cases, na.rm=T)-75, label = "MCO", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-05-04")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-05-04")+2, y = max(data_plot$new_cases, na.rm=T)-75, label = "CMCO", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-06-10")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-06-10")+2, y = max(data_plot$new_cases, na.rm=T)-75, label = "RMCO1", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-09-01")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-09-01")+2, y = max(data_plot$new_cases, na.rm=T)-75, label = "RMCO2", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-09-26")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-09-26")+2, y = max(data_plot$new_cases, na.rm=T)-75, label = "Sabah\nElection", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-12-07")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-12-07")+2, y = max(data_plot$new_cases, na.rm=T)-75, label = "Free\nTravel", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2021-01-13")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2021-01-13")+2, y = max(data_plot$new_cases, na.rm=T)-75, label = "MCO2", angle = 90) +
  labs(x = "Date", y = "New cases", title = paste0("New cases by ", format(sel_date, "%d %B %Y"))) +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1))# +
  theme_light()
plot1
ggplotly(plot1)
```

### Total cases, hover/tap on the graph for details.
```{r}
# cumulative
plot2 = ggplot(data_plot) +
  geom_col(aes(x = date, y = total_cases), fill = "red", alpha = 0.7) +
  geom_line(aes(x = date, y = total_cases_predicted), color = "blue", alpha = 0.7) +
  scale_x_date(date_breaks = "28 day", date_labels = "%d/%m/%y") +
  geom_vline(xintercept = as.numeric(as.Date("2020-02-28")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-02-28")+2, y = max(data_plot$total_cases, na.rm=T)-75, label = "Tabligh", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-03-18")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-03-18")+2, y = max(data_plot$total_cases, na.rm=T)-75, label = "MCO", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-05-04")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-05-04")+2, y = max(data_plot$total_cases, na.rm=T)-75, label = "CMCO", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-06-10")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-06-10")+2, y = max(data_plot$total_cases, na.rm=T)-75, label = "RMCO1", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-09-01")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-09-01")+2, y = max(data_plot$total_cases, na.rm=T)-75, label = "RMCO2", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-09-26")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-09-26")+2, y = max(data_plot$total_cases, na.rm=T)-75, label = "Sabah\nElection", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-12-07")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-12-07")+2, y = max(data_plot$total_cases, na.rm=T)-75, label = "Free\nTravel", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2021-01-13")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2021-01-13")+2, y = max(data_plot$total_cases, na.rm=T)-75, label = "MCO2", angle = 90) +
  labs(x = "Date", y = "Cumulative cases", title = paste0("Total cases by ", format(sel_date, "%d %B %Y"))) +
  theme_light()
plot2
ggplotly(plot2)
```

Row
---

### **Significant factors:** 1. Days since Tabligh gathering (`r paste0("Rate ratio: ", round(model1_rr[2, "estimate"],2), " [95% CI: ", round(model1_rr[2, "conf.low"],2), ", ", round(model1_rr[2, "conf.high"],2), "]")` per day), 2. Days since movement control order (`r paste0("Rate ratio: ", round(model1_rr[3, "estimate"],2), " [95% CI: ", round(model1_rr[3, "conf.low"],2), ", ", round(model1_rr[3, "conf.high"],2), "]")` per day), 3. Days since conditional movement control order (`r paste0("Rate ratio: ", round(model1_rr[4, "estimate"],2), " [95% CI: ", round(model1_rr[4, "conf.low"],2), ", ", round(model1_rr[4, "conf.high"],2), "]")` per day), 4. Days since recovery movement control order (`r paste0("Rate ratio: ", round(model1_rr[5, "estimate"],2), " [95% CI: ", round(model1_rr[5, "conf.low"],2), ", ", round(model1_rr[5, "conf.high"],2), "]")` per day) based on a quasipoisson regression model. {.value-box}
```{r}
valueBox(
  value = "",
)
```

Deaths
======

Row
---

### new deaths today  {.value-box}
```{r}
valueBox(
  value = data_plot1[data_plot1$date == date_now, "new_deaths"],
  #icon = "fa-area-chart",
  color = "red"
)
```

### new deaths expected tomorrow  {.value-box}
```{r}
valueBox(
  value = data_plot1[data_plot1$date == date_now + 1, "new_deaths_predicted"],
  #icon = "fa-area-chart",
  color = "red"
)
```

### additional new deaths expected in 3 days  {.value-box}
```{r}
valueBox(
  value = data_plot1[data_plot1$date == date_now + 3, "total_deaths_predicted"] - data_plot1[data_plot1$date == date_now, "total_deaths_predicted"],
  #icon = "fa-area-chart",
  color = "red"
)
```

### additional new deaths expected in a week  {.value-box}
```{r}
valueBox(
  value = data_plot1[data_plot1$date == date_now + 7, "total_deaths_predicted"] - data_plot1[data_plot1$date == date_now, "total_deaths_predicted"],
  #icon = "fa-area-chart",
  color = "red"
)
```

### total deaths today  {.value-box}
```{r}
valueBox(
  value = data_plot1[data_plot1$date == date_now, "total_deaths"],
  #icon = "fa-area-chart",
  color = "red"
)
```

### total deaths expected in a week  {.value-box}
```{r}
valueBox(
  data_plot1[data_plot1$date == date_now + 7, "total_deaths_predicted"] - data_plot1[data_plot1$date == date_now, "total_deaths_predicted"] + data_plot1[data_plot1$date == date_now, "total_deaths"],
  #value = data_plot1[data_plot1$date == date_now + 7, "total_deaths_predicted"],
  #icon = "fa-area-chart",
  color = "red"
)
```

Row
---

### New deaths, hover/tap on the graph for details.
```{r}
# daily
plot3 = ggplot(data_plot1) +
  geom_col(aes(x = date, y = new_deaths), fill = "red", alpha = 0.7) +
  geom_line(aes(x = date, y = new_deaths_predicted), color = "blue", alpha = 0.7) +
  scale_x_date(date_breaks = "28 day", date_labels = "%d/%m/%y") +
  scale_y_continuous(limits = c(0, max(data_plot1$new_deaths, na.rm=T)+50)) +
  geom_vline(xintercept = as.numeric(as.Date("2020-02-28")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-02-28")+2, y = max(data_plot1$new_deaths, na.rm=T)+30, label = "Tabligh", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-03-18")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-03-18")+2, y = max(data_plot1$new_deaths, na.rm=T)+30, label = "MCO",angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-05-04")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-05-04")+2, y = max(data_plot1$new_deaths, na.rm=T)+30, label = "CMCO", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-06-10")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-06-10")+2, y = max(data_plot1$new_deaths, na.rm=T)+30, label = "RMCO1", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-09-01")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-09-01")+2, y = max(data_plot1$new_deaths, na.rm=T)+30, label = "RMCO2", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-09-26")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-09-26")+2, y = max(data_plot1$new_deaths, na.rm=T)+30, label = "Sabah\nElection", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-12-07")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-12-07")+2, y = max(data_plot1$new_deaths, na.rm=T)+30, label = "Free\nTravel", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2021-01-13")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2021-01-13")+2, y = max(data_plot1$new_deaths, na.rm=T)+30, label = "MCO2", angle = 90) +
  labs(x = "Date", y = "New deaths", title = paste0("New deaths by ", format(sel_date, "%d %B %Y"))) +
  theme_light()
plot3
ggplotly(plot3)
```

### Total deaths, hover/tap on the graph for details.
```{r}
# cumulative
plot4 = ggplot(data_plot1) +
  geom_col(aes(x = date, y = total_deaths), fill = "red", alpha = 0.7) +
  geom_line(aes(x = date, y = total_deaths_predicted), color = "blue", alpha = 0.7) +
  scale_x_date(date_breaks = "28 day", date_labels = "%d/%m/%y") +
  scale_y_continuous(limits = c(0, max(data_plot1$total_deaths, na.rm=T)+50)) +
  geom_vline(xintercept = as.numeric(as.Date("2020-02-28")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-02-28")+2, y = max(data_plot1$total_deaths, na.rm=T)+30, label = "Tabligh", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-03-18")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-03-18")+2, y = max(data_plot1$total_deaths, na.rm=T)+30, label = "MCO",angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-05-04")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-05-04")+2, y = max(data_plot1$total_deaths, na.rm=T)+30, label = "CMCO", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-06-10")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-06-10")+2, y = max(data_plot1$total_deaths, na.rm=T)+30, label = "RMCO1", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-09-01")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-09-01")+2, y = max(data_plot1$total_deaths, na.rm=T)+30, label = "RMCO2", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-09-26")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-09-26")+2, y = max(data_plot1$total_deaths, na.rm=T)+30, label = "Sabah\nElection", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2020-12-07")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2020-12-07")+2, y = max(data_plot1$total_deaths, na.rm=T)+30, label = "Free\nTravel", angle = 90) +
  geom_vline(xintercept = as.numeric(as.Date("2021-01-13")), linetype = "dotted") +
  annotate(geom = "text", x = as.Date("2021-01-13")+2, y = max(data_plot1$total_deaths, na.rm=T)+30, label = "MCO2", angle = 90) +
  labs(x = "Date", y = "Cumulative deaths", title = paste0("Total deaths by ", format(sel_date, "%d %B %Y"))) +
  theme_light()
plot4
ggplotly(plot4)
```

Row
---

### **Significant factor:** 1. Number of cases in ICU (`r paste0("Rate ratio: ", round(model2_rr[2, "estimate"],2), " (95% CI: ", round(model2_rr[2, "conf.low"],2), ", ", round(model2_rr[2, "conf.high"],2), ")")` per increase in case in ICU). {.value-box}
```{r}
valueBox(
  value = "",
)
```

Cases by State
==============

Row
---

### New cases, hover/tap on the graph for details.
```{r}
# daily
plot5 = ggplot(data = data_plots) +
  geom_col(aes(x=date, y=new_cases, fill=state)) +
  scale_x_date(date_labels = "%d/%m/%y", limits = c(min(data_plots$date), NA)) +
  scale_fill_discrete(name="State") +
  labs(x="Date", y="New cases") + 
  theme_light()
plot5
ggplotly(plot5)
```

### Total cases, hover/tap on the graph for details.
```{r}
# cumulative
plot6 = ggplot(data = data_plots) +
  geom_col(aes(x=date, y=total_cases, fill=state)) +
  scale_x_date(date_labels = "%d/%m/%y") +
  scale_fill_discrete(name="State") +
  labs(x="Date", y="Cumulative cases") + 
  theme_light()
plot6
ggplotly(plot6)
```

Deaths by State
==============

Row
---

### New deaths, hover/tap on the graph for details.
```{r}
# daily
plot7 = ggplot(data = data_plots) +
  geom_col(aes(x=date, y=new_deaths, fill=state)) +
  scale_x_date(date_breaks = "28 day", date_labels = "%d/%m/%y") +
  scale_fill_discrete(name="State") +
  labs(x="Date", y="New deaths") + 
  theme_light()
plot7
ggplotly(plot7)
```

### Total deaths, hover/tap on the graph for details.
```{r}
# cumulative
plot8 = ggplot(data = data_plots) +
  geom_col(aes(x=date, y=total_deaths, fill=state)) +
  scale_x_date(date_breaks = "28 day", date_labels = "%d/%m/%y") +
  scale_fill_discrete(name="State") +
  labs(x="Date", y="Cumulative deaths") + 
  theme_light()
plot8
ggplotly(plot8)
```

Data and Our Team
=================

Row
---

###

**The data are sourced from:**

- Ourworldindata: https://covid.ourworldindata.org/data/ecdc/full_data.csv (up to 21/3/2020).
- Ministry of Health of Malaysia press statement: https://kpkesihatan.com/, http://www.moh.gov.my/index.php/pages/view/2019-ncov-wuhan-kenyataan-akhbar, https://www.facebook.com/kkmcprc/
- Our edited data can be accessed from https://data.world/wnarifin/covid-19-my/
- Last updated on `r format(Sys.Date(), "%d %B %Y")`.

**Epidemiology Modelling Team members, School of Medical Sciences, Universiti Sains Malaysia (USM):**

Department of Community Medicine:

- Assoc. Prof. Dr. Kamarul Imran Musa (Dr. KIM)
- Dr. Mohd Azmi B. Suliman
- Dr. Mohamad Zarudin B Mat Said
- Dr. Wira Alfatah B Ab Ayah @ Ab Aziz
- Dr. Che Muhammad Nur Hidayat B Che Nawi
- Dr. Afiqah Syamimi Bt Masrani
- Mr. Tengku Muhammad Hanis B Tengku Mokhtar

Biostatistics and Research Methodology Unit:

- [Dr. Wan Nor Arifin](https://wnarifin.github.io/)

Jabatan Kesihatan Negeri, Pahang:

- Dr. Sahrol Azmi

Pusat Data PPKT Kampus Kesihatan:

- Mohd Fadzali Bakar
- Mohd Faizal Abdul Manaf
- Ahmad Syakiren Mazalan 

**Disclaimer: We and USM are not responsible for any damage or inconvenience caused directly or indirectly by the use of this dashboard.**

We built this dashboard using

- RStudio
- Packages:

    - flexdashboard
    - ggplot2
    - plotly
    - broom
